\chapter*{Conclusion}

The analysis shows how an iterative process can be a solution to find the best parameters for a genetic algorithm implementation.
I found it hard to find a linear path to the solution, but the iterative process helped me to find the best parameters for the genetic algorithm. I learned how this kind of tuning process is very difficult to standardize due to the uniqueness of each problem. I am pround and happy of the results I got, especially regarding the framework created to test and then analyze the results. I think this was the key to succeed in these kinds of experimental projects.

The problem task suggested to run each test ten times and then perform the average of the results. I think this is a good approach to reduce the noise in the results, but I feel that for this algorithm implementation it wasn't enought. It looked like different executions gave too different results, and I think that running the tests more times could have helped to reduce the noise even more, allowing for more accurate results.

Despite the obtained results, I believe it would be interesting to approach the problem using a metaheuristic algorithm, such as the genetic algorithm itself, to solve the problem of finding the best parameters for the genetic algorithm. The iterative process is slow and requires a lot of manual work, while a metaheuristic algorithm could be more challenging and fun.

I really enjoyed working on this project, I think I learned a lot about genetic algorithms and how to tune them. The task problem was clear and having the test instances ready saved me plenty of time. This was my first experimental research project, and the freedom to choose any language and technology I believe was the essence of the project, even though it was a bit challenging to find the best tools to use and I was a little bit lost in the beginning. When looking for literature resources, I found the No Free Lunch Theorem \cite{NFL-theorem}, which left me slightly unsettled but gave me a different approach to the problem. I hope to have the opportunity to work on more projects like this in the future.


\addcontentsline{toc}{chapter}{Conclusion}
